Cluster Difference Imaging Photometric Survey. I. Light Curves of Stars in Open Clusters from TESS Sectors 6 \& 7

### 2019-08-10 paper_I/ms.tex
update intro
Co-authored-by: joshwinn <jwinn@astro.princeton.edu>

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}

@@

\title{
  CDIPS. I. Light Curves in Open Clusters from TESS Sectors 6 \& 7
}

@@

\begin{abstract}
  % Context
  The Transiting Exoplanet Survey Satellite (TESS) has begun to
  photometrically monitor almost every star cluster in the solar
  neighborhood.
  % Aims
  Using the TESS images, we have begun a Cluster Difference Imaging
  Photometric Survey (CDIPS), in which we are focusing on stars that
  are candidate cluster members, or else show other indications of
  youth relative to field stars.
  %
  Our aims are to discover giant transiting planets with known
  ages, and also to provide a light curve sample fit for studies in
  stellar astrophysics.
  %
  For this work, we made \numberlcs light curves of candidate young
  stars, across \numberclusters distinct clusters.  Each light curve
  represents between 20 and 25 days of observations of a star brighter
  than $G_{Rp}=16$, sampled at 30 minute cadence.
  % Methods
  We describe the image subtraction and time-series analysis
  techniques we used to create the light curves, which have noise
  properties that agree with theoretical expectations.
  %
  We also comment on the possible utility of the light curve sample for
  studies of stellar rotation evolution, and binary eccentricity
  damping.
  %
  The light curves, which cover about one sixth of the galactic plane,
  are available at \stscilink.
\end{abstract}

@@

Each of the $\sim$3000 known star clusters of the Milky Way is a gift
to astrophysics, providing a sample of stars that vary widely in mass
but all have approximately the same age and composition.  Time-series
photometry of clusters has many applications. By measuring rotation
periods over a range of ages, we can study the angular momentum
evolution of stars and improve our ability to determine stellar ages
through gyrochronology \citep[{\it
e.g.},][]{skumanich_time_1972,barnes_color-period_2015,meibom_spin-down_2015,curtis_tess_2019}.
By measuring eccenticities as a function age, we can study the tidal
circularization of binaries
\citep{meibom_robust_2005,milliman_wiyn_2014,price-whelan_binary_2018}.
The detection of eclipsing binaries (EBs) can also lead to the precise
determination of the absolute dimensions of the stars and stringent
tests of stellar-evolutionary models
\citep{luhman_formation_2012,stassun_review_2014,kraus_mass-radius_2015}.
Finally, transiting exoplanets discovered in clusters can shed light
on the timescales for processes in planet formation, evolution, and
migration
\citep[][]{Fortney_et_al_2007,Mann_K2_33b_2016,David_et_al_2017}, as
well as on the effects of metallicity
\citep[][]{fischer_planet-metallicity_2005,petigura_metallicity_2018}.

The TESS mission \citep{ricker_transiting_2015}, though designed for
other purposes, holds the promise to deliver the most homogeneous and
comprehensive cluster photometric survey in history.  More
quantitatively, the \citet{Kharchenko_et_al_2013} cluster member
database indicates that $\approx 2\times10^5$ open cluster members
brighter than $T=16$ will be observed in the full-frame images over
the first two years of TESS observations.  This count includes the
``most probable'' members within each cluster's radius, which by the
\citet{kharchenko_global_2012} definition requires the independent
kinematic, photometric, and spatial membership probabilities to each
exceed 61\%.  The actual number of stars in clusters is larger, as the
membership catalogs are not yet complete, even at these
relatively bright magnitudes \citep[{\it
e.g.},][]{roser_nine_RSG_2016,cantat-gaudin_gaia_2018,cantat-gaudin_newOCs_2019}.

A major barrier to deriving precise photometry from the TESS images is
the relatively poor angular resolution ($\approx 21''$ per pixel).
Almost all clusters are within 10 degrees of the Galactic plane. The
problems with crowding and complex backgrounds become so severe
that the TESS Candidate Target List deprioritizes
2-minute targets at galactic latitudes less than $\approx
15^\circ$~\citep{stassun_TIC_2018,stassun_TIC8_2019}. This includes
almost all star clusters\footnote{TICv8 updated the cutoff to
$10^\circ$. During the first year of TESS observations, the number
used when selecting 2-minute target stars was $15^\circ$.}.  This decision
was made because
the large pixel size and the high stellar surface density make
aperture photometry unreliable.  However, it means that most
stars in clusters, though capable of yielding a cornucopia of science,
will go unprocessed by the official {\it TESS} data reduction
pipeline.

One way to quantify the blending problem is to determine what fraction of
the total flux in a photometric aperture is contributed by a
particular target star. Aperture photometry is reliable when this
fraction is close to unity.  
Difference imaging \citep{Alard_Lupton_1998,miller_optimal_2008}, in
our group's experience, can be viable down to crowding fractions
of perhaps 1 part in 10.
Based on the
\citet{Kharchenko_et_al_2013} cluster membership data, and the density
of background stars, we calculated that the median dilution fraction of
cluster stars with $T< 16$ is 0.13, for an aperture radius of 2 pixels.
For at least $\sim$$10^5$ cluster stars then, it seems that difference
imaging may be worthwhile.

Difference imaging avoids the primary effects of blending through
forced-aperture photometry.  In this method,
the positions of stars are predicted from an astrometric
solution, and the reference fluxes are determined from a calibrated
catalog magnitude-to-flux relation.  The deviation from the reference
flux is measured on the difference image.  Assuming that only a single
source causes the variability, blended neighbor stars only act to
increase Poisson noise, and do not ``dilute'' transit signals, down to
the angular resolution of the source catalog used to determine the
reference flux.  This is a major benefit of performing image
subtraction in crowded fields.

We have therefore begun to apply difference imaging to the
TESS images, with a focus on any star that could be a member
of a coeval group.  We
are also including some stars that we suspect are young due to
combined photometric and astrometric indicators.
A major motivation for
this effort is to discover giant transiting planets with known ages.
The focus of the present study however is to describe our
methods, and to produce a general-purpose
dataset applicable for studies both in exoplanetary and also
stellar astrophysics.

For the remainder of this work, we adopt the term ``star cluster'', or
simply cluster, to refer to a coeval group of stars.  This includes
the open clusters of yore, as well as the moving groups and stellar
associations that have been discovered since the late
1990s \citep{zuckerman_young_2004}.

The plan of action is as follows. In \S~\ref{sec:starselection}, we
describe how target stars were selected. \S~\ref{sec:method} presents
the photometric and image processing methods we used to produce light
curves for these stars.  The statistical properties of the \numberlcs
light curves from TESS Sectors 6 and 7 are then summarized
(\S~\ref{subsec:lcstatistics}).  An example of how we might use the
data to identify pulsating stars, eclipsing binaries, and transiting
planets is then given (\S~\ref{subsec:identifying_variability}).
\S~\ref{sec:conclusion} summarizes our findings, and discusses them
with an eye towards the studies we hope this and subsequent rounds of
data processing will enable.

@@

\subsubsection{Photometric reference frame construction \& reference flux measurement}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO
% * update keywords per new AAS format

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt,twocolumn,tighten]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}

@@

\title{
  Cluster Difference Imaging Photometric Survey. I.
  Light Curves of Stars in Open Clusters from TESS Sectors 6 \& 7
}

@@

\begin{abstract}
  % Context
  The Transiting Exoplanet Survey Satellite (TESS) is providing
  precise time-series photometry for hundreds of nearby star clusters.
  % Aims
  Using the TESS images, we have begun a Cluster Difference Imaging
  Photometric Survey (CDIPS), in which we are focusing on stars that
  are candidate cluster members, or else show other indications of
  youth.
  %
  Our aims are to discover giant transiting planets with known ages,
  and to provide light curves suitable for studies in stellar
  astrophysics.
  %
  For this work, we made \numberlcs light curves of candidate young
  stars, across \numberclusters distinct clusters.  Each light curve
  represents between 20 and 25 days of observations of a star brighter
  than $G_{Rp}=16$, with 30-minute sampling.
  % Methods
  We describe the image subtraction and time-series analysis
  techniques we used to create the light curves, which have noise
  properties that agree with theoretical expectations.
  %
  We also comment on the possible utility of the light curve sample
  for studies of stellar rotation evolution, and binary eccentricity
  damping.
  %
  The light curves, which cover about one sixth of the galactic plane,
  are available at \stscilink.
\end{abstract}

@@

Each of the several thousand star clusters of the Milky Way is a
gift to astrophysics, providing a sample of stars that vary widely in
mass but all have approximately the same age and composition.
Time-series photometry of clusters has many applications. By measuring
rotation periods over a range of ages, we can study the angular
momentum evolution of stars and improve our ability to determine
stellar ages through gyrochronology \citep[{\it
e.g.},][]{skumanich_time_1972,barnes_color-period_2015,meibom_spin-down_2015,curtis_tess_2019}.
By measuring the eccentricity distribution of binary stars as a
function of age, we can study the tidal circularization process
\citep{meibom_robust_2005,milliman_wiyn_2014,price-whelan_binary_2018}.
The detection of eclipsing binaries (EBs) can also lead to the precise
determination of the absolute dimensions of the stars and stringent
tests of stellar-evolutionary models
\citep{luhman_formation_2012,stassun_review_2014,kraus_mass-radius_2015}.
Finally, transiting exoplanets discovered in clusters can shed light
on the timescales for processes in planet formation, evolution, and
migration
\citep[][]{Fortney_et_al_2007,Mann_K2_33b_2016,David_et_al_2017}, as
well as on the effects of metallicity
\citep[][]{fischer_planet-metallicity_2005,petigura_metallicity_2018}.

The Transiting Exoplanet Survey Satellite (TESS,
\citealt{ricker_transiting_2015}) holds the promise to deliver the
most homogeneous and comprehensive cluster photometric survey in
history.  Based on the cluster membership data of
\citet{Kharchenko_et_al_2013}, approximately $2\times10^5$ open
cluster members brighter than $T=16$ will be observed in the
full-frame images over the first two years of TESS observations.  This
count includes the ``most probable'' members, which by the
\citet{kharchenko_global_2012} definition requires the independent
kinematic, photometric, and spatial membership probabilities to each
exceed 61\%.  The actual number of stars in clusters is larger, as the
membership catalogs are not yet complete, even at these relatively
bright magnitudes \citep[{\it
e.g.},][]{roser_nine_RSG_2016,cantat-gaudin_gaia_2018,cantat-gaudin_newOCs_2019,kounkel_untangling_2019,sim_open_2019}.

A major barrier to deriving precise photometry from the TESS images is
the relatively poor angular resolution ($21''$ per pixel).  Almost all
clusters are within 10 degrees of the Galactic plane. The problems
with crowding and complex backgrounds are so severe that the TESS
Candidate Target List deprioritizes 2-minute targets within $15^\circ$
of the galactic plane\footnote{During the first year of TESS
observations, objects within $15^\circ$ of the galactic plane were
deprioritized. Starting in the second year of observations,
the cutoff was changed to
objects within $10^\circ$ of the galactic plane
\citep{stassun_TIC8_2019}.}~\citep{stassun_TIC_2018,stassun_TIC8_2019}.
This decision was made because the large pixel size and the high
stellar surface density make aperture photometry unreliable.  However,
it implies that most stars in clusters, though capable of yielding a
cornucopia of science, will go unprocessed by the official {\it TESS}
data reduction pipeline.

One way to quantify the blending problem is to determine what fraction
of the total flux in a photometric aperture is contributed by a
particular target star. Aperture photometry is reliable when this
fraction is close to unity.  Difference imaging
\citep{Alard_Lupton_1998,miller_optimal_2008}, in our group's
experience, can be viable down to crowding fractions of 10\%.  Based
on the \citet{Kharchenko_et_al_2013} cluster membership data, and the
density of background stars, we calculated that the median dilution
fraction is 0.13, for cluster stars with $T< 16$ and an aperture
radius of 2 pixels.  Thus, for at least $\sim$$10^5$ cluster stars,
difference imaging may be advantageous.

Difference imaging avoids the primary effects of blending through
forced-aperture photometry.  In this method,
the pixel coordinates
of stars are calculated from an astrometric
solution, and the reference fluxes are determined from a calibrated
catalog magnitude-to-flux relation (see \S~\ref{subsubsec:photref}).
The deviation from the reference
flux is measured on the difference image.  Assuming that only a single
source is variable, blended neighbor stars only act to
increase Poisson noise, down to
the angular resolution of the source catalog used to determine the
reference flux.  This is a major benefit of performing image
subtraction in crowded fields.

We have therefore begun to apply difference imaging to the
TESS images, with a focus on any star that could be a member
of a group.  We
are also including some stars that we suspect are young due to
combined photometric and astrometric indicators.
A major motivation for
this effort is to discover giant transiting planets with known ages.
The focus of the present study however is to describe our
methods, and to produce a general-purpose
dataset applicable for studies both in exoplanetary and also
stellar astrophysics.

For the remainder of this work, we adopt the term ``star cluster'', or
simply cluster, to refer to a coeval group of stars.  This includes
open clusters, as well as the moving groups and stellar associations
that have been discovered since the late 1990s
\citep{zuckerman_young_2004}.

The plan of action is as follows. In \S~\ref{sec:starselection}, we
describe how target stars were selected. We then present the
photometric and image processing methods we used to produce light
curves for these stars (\S~\ref{sec:method}).  In TESS Sectors 6 and
7, these methods yielded \numberlcs light curves, and in
\S~\ref{subsec:lcstatistics} we summarize their statistical properties.
An example of how we might use the data to identify pulsating stars,
eclipsing binaries, and transiting planets is then given
(\S~\ref{subsec:identifying_variability}).  We close in
\S~\ref{sec:conclusion} by summarizing our findings, and discuss them
with an eye towards the studies we hope this and further data
processing will enable.

@@

\subsubsection{Photometric reference frame construction \& reference flux measurement}
\label{subsubsec:photref}
